### 2025-07-07 conteudo/methods/methods.tex
Add todo note about a typo
update with the papers updates and revisions

Given the tool evaluations demonstrating the tool's capabilities of finding duplications, 
we proceeded to understand if the duplications found can be mitigated and the kernel 
subsystem's maintainers' opinions about the duplications and their respective mitigations. 
To accomplish this objective, we proposed a participant-observation experiment, 
where we, people who have never contributed to the Linux kernel, act as contributors to 
the AMD Display driver and collect artifacts of our experience in the process.

We executed the tool in the AMD Display driver codebase. 
We manually analyzed the biggest duplications regarding the number of lines and raised them 
to two duplicated function pairs that we judged interesting to try to mitigate. We have 
chosen to approach the biggest duplications as proof of concept so that we can make 
significant impacts. The two duplicated function pairs chosen are:

\begin{itemize}
\item \textbf{Duplicated function pair 1}
\begin{itemize}
\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c}.
\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
\end{itemize}
\item \textbf{Duplicated function pair 2}
%\begin{itemize}
\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce110/command\_table\_helper\_dce110.c}.
\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce60/command\_table\_helper\_dce60.c}.
%\end{itemize}
\end{itemize}

Given the two duplicated function pairs, we observed that the files in the functions 
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic approach, for the duplicated 
function pairs that had successfully mitigated the duplication, we sent patches to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

@@

In C programming, communication between source files is achieved by creating header files that specify 
libraries \citep{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
by consolidating duplicated code into a single library, which would replace instances of 
duplication across the codebase.

For each duplicated function pair, for each function in the pair, we used the \textit{function information} functionality 
(described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

@@

A core dynamic of the Free Software Development course at the University of São Paulo is to 
have students make a practical contribution to a free software project. In the first semester of 
2025, the professor teaching the course, who is also the advisor of this work, identified an 
opportunity to leverage the ArKanjo tool within this existing pedagogical structure. As a 
contribution option, the professor proposed that students mitigate code duplications found in the 
Linux kernel. This initiative was driven by the professor as a practical course activity, and 
the author of this thesis was not involved in using the tool to find duplications nor in developing 
the approaches to mitigate them. The tasks were carried out entirely by the course staff 
and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
one to three. To assist students in this task, the professor and teaching assistants prepare multiple 
alternatives for contributing to free software projects, including simpler contributions such as 
updating documentation or fixing code style issues. In this context, two new options related to the 
ArKanjo tool were proposed alongside the others.

For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem, 
creating a list of duplicated function pairs found with a 90\% similarity threshold. The students' 
task was to refactor the code to mitigate these specific duplications and then navigate the process 
of submitting their patches to the IIO maintainers for review.

For the second alternative, students were offered an experience similar to our initial participant 
observation. In this model, the students were responsible for the entire workflow: executing the 
ArKanjo tool on the AMD display driver, independently analyzing the results to identify a duplication 
to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.

Of the students who chose to work on the ArKanjo-related tasks, 25 (20 undergraduate and 5 graduate) 
formed 11 groups to pursue the first alternative. One graduate student opted for the second alternative, 
forming a single group. Notably, none of these students had prior experience contributing to the 
Linux kernel, making this their first attempt to submit a patch as newcomers.

The groups were requested to document their experience and approaches to refactoring the 
duplications and the experience of submitting a patch to the Linux kernel in blogs. We analyzed 
these blogs to understand the patterns of refactoring used for mitigating the duplications, the 
experience of sending the patches, and the opinions of the subsystem maintainers on the mitigations.

The teaching assistant and the student who executed the ArKanjo tool to explore duplications were 
asked to answer some questions related to their experience using the tool and what could be improved. 
The survey with the users' answers can be found in Appendix \ref{app:survey}.

In the second stage of the Free Software Development course, we propose that students contribute 
directly to the ArKanjo tool development without help from the tool creator to verify whether the 
tool can be free software to which multiple people can contribute. The students approached the tool’s 
issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
groups worked in direct contributions.

>>>>

To investigate whether we can use the ArKanjo tool to find ways to contribute to the 
Linux kernel, we conducted a participant observation experiment in the second semester 
of 2024, where we acted as first-time contributors to the AMD Display driver, collecting 
artifacts of our experience in the process.

In the AMD Display driver codebase, we executed the ArKanjo tool. 
We manually analyzed the biggest duplications regarding the number of lines and found 
two duplicated function pairs that we judged interesting to try to mitigate. We 
have chosen to approach the biggest duplications as proof of concept so that we 
can make significant impacts.

The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}. The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file 
\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.

Given the two duplicated function pairs, we observed that the files in the functions 
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic strategy, for the duplicated 
function pairs that had successfully mitigated the duplication, we sent patches to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

@@

In C programming, communication between source files is achieved by creating header files that specify 
libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
by consolidating duplicated code into a single library, which would replace instances of 
duplication across the codebase.

For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

@@

A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to 
have students make a practical contribution to a free software project. In the first semester of 
2025, the professor teaching the course, who is also the advisor of this work, identified an 
opportunity to leverage the ArKanjo tool within this existing pedagogical structure. 
As a contribution option, the professor proposed that students mitigate code duplications found 
in the Linux kernel. The professor drove this initiative as a practical course activity, 
and the author of this thesis was not involved in using the tool to identify duplications or 
in developing approaches to mitigate them. The tasks were carried out entirely by the course 
staff and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of
two or three members. To assist students in this task, the professor and teaching assistants prepare multiple
alternatives for contributing to free software projects, including simpler contributions such as
updating documentation or fixing code style issues. In this context, two new options related to the
\texttt{anonymous-tool} were proposed alongside the others.

For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, 
creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other 
with a 90\% similarity threshold. Then, the following process happened to both lists:

\begin{enumerate}
    \item The teaching assistant filtered the lists to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
    \item The teacher assistant removed duplications that were too short and would not represent a significant change for a patch application;
    \item The teacher assistant manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries. This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
\end{enumerate}

The students' task was to select a recommended entry from the list, refactor the code to eliminate 
these specific duplications, and then submit their patches to the IIO maintainers for review.

For the second alternative, students were offered an experience similar to our initial participant
observation. In this model, the students were responsible for the entire workflow: executing the
ArKanjo on the AMD display driver, independently analyzing the results to identify a duplication
to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.

Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 
23 (16 undergraduate and 7 graduate)
formed 11 groups to pursue the first alternative. One graduate student also opted for the second 
alternative. Notably, none of these students had prior experience contributing to the Linux kernel, 
making this their first attempt to submit a patch as newcomers to the project.

The groups were requested to document their experiences and approaches to refactoring the duplications,
as well as their experiences submitting a patch to the Linux kernel, in blogs. We analyzed these 
blogs to understand the refactoring patterns used to mitigate duplications, the experience of 
sending patches, and the opinions of subsystem maintainers on the mitigations.

### 2025-07-07 conteudo/results/participant.tex
update with the papers updates and revisions

For our participant observation, we attempted to mitigate two code 
duplications found by the ArKanjo tool in the AMD Display driver by applying 
our systematic approach presented in \ref{subsubsec:systematic}. 
As presented in section \ref{sec:meteth},
we opted to apply the systematic approach in the functions
\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
\textit{dc/gpio} and \text{dc/bios}, respectively.

Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. 
Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. In the next sections, 
we present our experience and learning for each mitigation.

@@

In the systematic approach, we found that the \textit{offset\_to\_id} function 
exists in code 10 files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to 
each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

We reached a point with a scratch on how to mitigate the duplications found, 
but we could not complete the refactoring on a state capable of submitting as 
a patch to the driver. We found that the configuration files on the AMD Display 
driver are designed to be imported at compilation time using \textit{\#define} macros. 
The configuration files design choice makes the refactoring of duplications on 
generic approaches trickier, as refactoring code that depends on these files 
requires significant refactoring on the configuration files design and deep 
knowledge of the codebase, which we as first-timer contributors, do not have. 
Thus, we opted not to continue investigating this mitigation.

@@

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom}
function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring work was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus, 
we did not face the same issues during the first mitigation. Refactoring the code 
to mitigate duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver’s maintainers as a patch, documenting our findings and experiences 
on the process.

@@

The experience of submitting a patch was not as we expected. The process took 7 months 
and 16 days, which is significantly slower than expected. We initially submitted the 
patch on August 9th, 2024, but we did not receive replies. Thus, we had to resend it 
on October 9th, 2024. We received an initial response on November 3rd, 2024, asking for 
minimal changes; we submitted a second version with the changes request on November 16th, 2024. 
We did not receive replies at first and need to resend the patch on February 7th, 2025. 
We received a reply on February 11th, 2025, asking for additional changes. We sent a third 
version on February 24th, 2025, and the patch was accepted and integrated into the kernel 
codebase on February 25th, 2025. Figure \ref{fig:timeline} illustrates the timeline of 
the patch submission.

We discussed with our point of contact at the AMD Display driver to understand the reason 
for the slow process of submitting a patch. We learned that some of the code in the driver 
is shared across all operating systems that support the implemented GPUs. This fact creates 
a complex process within AMD to format and submit changes across the supported systems while 
simultaneously implementing measures to mitigate errors. Additionally, we comprehended that 
AMD Display driver developers can not view duplicated code negatively, as we understand it 
is a good practice in software engineering. One example shared with us is that duplicated 
code enhances the independence of GPU driver code, allowing developers to make changes to a 
specific GPU without needing to test compatibility with others. 
This approach helps save significant time and effort.

Regarding the patch changes requested for the second version, the maintainers only asked 
for minor changes to align with code style, license correct use, and best practices that 
were initially unknown to use. For the change requested on the third version, the maintainers 
asked us to move the functions created in the generic library to an existing file instead of 
creating a new one. More details of the patch can be found in Appendix \ref{app:feedback}, 
and the patch submitted in the Linux codebase can be seen here:
\url{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}.

>>>>

For our participant observation, we attempted to mitigate two code 
duplications found by the ArKanjo tool in the AMD Display driver by applying 
our systematic approach presented in \ref{subsubsec:systematic}. 
As presented in section \ref{sec:meteth},
we opted to apply the systematic approach in the mitigations.

Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. In the next sections, 
we present our experience and learning for each mitigation.

@@

In the systematic approach, we found that the \textit{offset\_to\_id} function
exists in 10 code files, enclosing multiple GPU architectures. The duplicated
functions are not exactly equal, as some minimal specific logic is applied to 
each GPU architecture. To address this issue, we used the extract method to
break the function into smaller parts and the parameterized method to address
the specific logic for the GPUs.

We reached a point where we had a scratch on how to mitigate the duplications found, 
but we could not complete the refactoring to a state capable of submitting as a 
patch to the driver. We found that the configuration files on the AMD Display 
driver are designed to be imported at compilation time using \textit{\#define} macros. 
The configuration files' design choice makes refactoring duplications on generic 
approaches trickier, as refactoring code that depends on these files requires 
significant refactoring of the configuration files' design and deep knowledge 
of the codebase, which we, as first-time contributors, do not have. Thus, we opted 
not to continue investigating this mitigation.

@@

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function 
exists in five code files, and two other functions are duplicated
in those five code files. All functions across the files were exactly equal,
so there was no need to apply the refactoring methods presented in the literature.
The refactoring was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus,
we did not face the same issues during the first mitigation. Refactoring the code
to mitigate duplications reached a point we judged was good enough to submit as a
patch to contribute to the AMD Display driver code quality. Thus, we moved to send
the refactoring to the driver maintainers as a patch, documenting our findings and experiences
on the process.

@@

The experience of submitting a patch was not as we expected. The process took 7 
months and 16 days, significantly slower than expected. We initially submitted 
the patch on August 9, 2024, but we did not receive replies. Thus, we had to 
resend it on October 9, 2024. We received an initial response on November 3, 2024, 
requesting minimal changes. We submitted a second version with the requested 
changes on November 16, 2024. We did not receive replies initially and will 
need to resend the patch on February 7, 2025. We received a reply on 
February 11, 2025, asking for additional changes. We sent a third version on 
February 24, 2025, and the patch was accepted and integrated into the kernel 
codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the 
timeline of the patch submission.

After the initial reply from the maintainers, we observed that the review process 
became more responsive. We hypothesize that once a patch receives initial attention, 
it is more closely tracked by maintainers. Another contributing factor to the initial 
delay could be the maintainers' workload at the time of the first submission.

We discussed the Display driver with our point of contact at AMD to understand the 
reason for the slow patch submission process. We learned that some of the code in 
the driver is shared across all operating systems that support the implemented GPUs. 
This fact creates a complex process within AMD to format and submit changes across 
the supported systems while simultaneously implementing measures to mitigate errors. 
Additionally, we comprehended that AMD Display driver developers can not view 
duplicated code negatively, as we understand it is a good practice in software engineering. 
One example shared with us is that duplicated code enhances the independence of 
GPU driver code, allowing developers to make changes to a specific GPU without 
needing to test compatibility with others. This approach helps save a 
significant amount of time and effort.

Regarding the patch changes requested for the second version, the maintainers 
only asked for minor changes to align with code style, correct license use, 
and best practices that were initially unknown. For the change requested on 
the third version, the maintainers asked us to move the functions created in 
the generic library to an existing file instead of creating a new one.
\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

### 2025-07-07 conteudo/results/students.tex
update with the papers updates and revisions

For the students who worked in the IIO subsystem, the teaching assistant gave them duplicated 
function pairs 
found by the tool, and their responsibility was to think of how to mitigate the duplication 
and send the patch to the maintainers. The student that worked in the AMD Display driver 
have the additional work to use the tool to find the duplications.

@@

Table \ref{tab:stu} summarizes students' groups' mitigation approaches. Seven groups applied 
the parameterized refactor method, 
and two applied the extractor method to their approaches to mitigate the duplications. 
The other three groups do not use any refactor methods presented in the literature.  
Five groups got their patches accepted and merged in Linux; four groups got their patches rejected, 
and one group is still working on the patch submission, being in the fifth version of the patch at 
the time this document was written. Two groups decided not to continue sending the patches after 
they made initial versions.

Groups one and four have worked in similar duplications, where two functions return a 
boolean mirroring the results. That means when the first function returns true, the second 
function returns false, and when the first function returns false, the second function 
returns false. Group one had a simple approach, and the maintainers approved the changes quickly. 
Group four attempted to mitigate the duplication with macros, a method found in the code file 
they were working on. The maintainers disagreed with the proposed approach, arguing that the 
proposed changes decrease the code readability, and asked for an approach similar to group one. 
Group four submitted the patch after coding the approach proposed by the maintainers.

Groups three, six, and ten created a struct to encapsulate the parameters and mitigate duplication 
when applying the parameterized method. Group three sent the mitigation to the maintainers and 
approved it without significant comments. In contrast, the patch from group ten was rejected by 
the same maintainer who approved the patch from group three. The maintainer justification to reject 
the patch is that group ten refactored code files that has simple context, therefore adding a layer 
of abstraction to remove a duplication deacrease the code readability without clear gains on 
the codebase quality. Figure \ref{fig:group_10_reply} shows the maintainers's feedback for group ten.

\begin{figure}
\includegraphics[scale=0.8]{group_10_reply}
\caption[Maintainers' Feedback on Group Ten's Patch]{
Maintainers' Feedback on Group Ten's Patch
\footnote{Image obtained from Linux Lore, which is publicly provided at:
\url{https://lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/\#u}}
}
\label{fig:group_10_reply}
\end{figure}

Groups seven and eight used the extract method to approach the mitigations. Group eight sent the 
mitigations to the maintainers, and the code changes were approved, with minor requested changes 
in code style and smaller errors fixed. For group seven mitigation, the maintainers reject their 
patch because the code changes add layers of abstractions to the code, making it more complex and 
challenging to read. The maintainers proposed that group seven remove the duplicated functions and 
insert the duplicated code multiple times where the functions were being called. This suggestion 
may be bad practice for code quality, as it adds more complexity and responsibility to a bigger 
function; it is viewed as the best choice for removing layers of abstraction and making the 
code easier to read. Figure \ref{fig:group_7_reply} shows the maintainers's feedback for group seven.

\begin{figure}
\includegraphics[scale=0.8]{group_7_reply}
\caption[Maintainers' Feedback on Group Seven's Patch]{
Maintainers' Feedback on Group Seven's Patch
\footnote{Image obtained from Linux Lore, which is publicly provided at:
\url{https://lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}}
}
\label{fig:group_7_reply}
\end{figure}

Groups nine and eleven faced similar issues to those in Mitigation 1 on our participant observation 
experiment, where configuration files designed as macros complicated the approach to the mitigations. 
Group nine applied the parameterized method to not need to know the configurations in the 
generic function, which is reasonable as the duplication approach only depends on one configuration 
value. Group eleven worked with multiple configuration values, and to refactor the duplication, 
they proposed to create a bigger macro that encapsulates the other macros, which is a clever 
solution but adds abstraction and complexity to macros, which makes the code harder to debug and read.
We do not get the feedback of the maintainers on the group eleven approach to know their opinions.

Group twelve is the student who executed the tool and passed through the process of finding 
duplications. The student found a duplication in the driver that happened across eighteen code files, 
creating a patch to remove four hundred lines of duplicated code. The student sent the mitigations 
to the maintainers, and their feedback pointed to fixing the warning in the code without arguing 
about removing the duplications.

Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that 
people without previous experience in Linux kernel development could approach duplications found 
by the ArKanjo tool and become contributors to the kernel. We observed that not all duplications 
are viewed as code with bad quality, with the maintainers analyzing the trade-off between the 
purpose of the duplicated code, code readability, and the levels of abstraction added to 
address the duplications.

>>>>

@@

Table \ref{tab:stu} summarizes the student groups' mitigation approaches. It contains 
the information about the chosen subsystem and the similarity threshold for groups that 
opted for the first approach. Each entry also has the patch status, refactoring method, 
and code diff corresponding to the final patch versions before it was dropped or accepted 
into the code base. The code diff for a patchset corresponds to the sum of added and removed 
lines of each of its patches.

Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications. Four patches do not contain a refactoring technique defined in the literature (undefined). Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.

Groups 1 and 4 have worked in similar duplications, where two functions are involved: one to determine if a register is volatile and the other to determine if it is writable. These are two functions that return opposite Boolean results. Group 1 initially addressed the problem using the extract method, while Group 2 used macros. After feedback from maintainers, both solutions converged to keep one of the functions while changing the content of the other to return a negated query of the former function. However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted. The reason is that Group 1 has chosen to apply the refactoring to a device that contains a register that is both read-only and non-volatile. Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.

Groups 3, 6, and 10 created a struct to encapsulate the parameters and mitigate duplication when applying the parameterized method. The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3. The maintainer's justification for rejecting the patch from Group 6 is that it proposes to fix duplication while adding more lines than removing, and is not convincing enough in other aspects, such as readability. The reason for rejecting the patch from Group 10 is that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreases code readability without clear gains in codebase quality. The maintainers' feedback for Group 10 was:
\begin{quote}

\textit{``A few comments inline, but overall I'm not sure the code reduction is sufficient to cover the resulting loss of readability. Sometimes a switch is simply clearer than a partial look-up table.''}

\textit{``So nice idea, but I'm not seeing it as being a good move for this particular code. I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high-performance path, so we care a lot more about readability here.''}
\end{quote}

Groups 7 and 8 used the extract method to approach the mitigations. Group 8 sent the mitigations to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed. For Group 7 mitigation, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function. After the feedback cycle, maintainers accepted the patch in version 3.

Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt. The patch is still under development, currently at version 6, and has been separated into a 4-patch patch set, where half of them have been applied, and the remaining entries are receiving suggestions. Fortunately, maintainers are receptive and provide constant feedback; the contribution is expected to be fully merged if the students do not abandon the current work.

Group 11 is the student who executed the tool and completed the process of identifying duplications. The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code. The student sent the mitigations to the maintainers, and their feedback pointed to fixing the warning in the code without arguing about removing the duplications. After addressing the issue, the patch was applied. This student also contributed to IIO, tackling a code duplication with inline method refactoring. This one, however, was received with more skepticism by the maintainer, and the student dropped the contribution because he did not think the feedback instructions were clear enough.

Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel. We observed that not all duplications
are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.

\subsection{Statistics of the contributions}

\todo[inline]{Is it ok to add it here in the text? It was Spessoto that writed this section entirely and I reviewed it}

The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively 
requested the insertion of 379 lines and the deletion of 605 lines in the subsystem, 
representing an average of 34.45 lines added and 55 lines removed per contribution.

Considering the sum of fully merged contributions, we have a code diff value of +51/-143, 
with an average of +12.75/-35.75. This ratio is better than dropped contributions, which 
have a total of +179/-219 with an average of +29.83/-36.5.

It is also possible to notice a difference based on the similarity threshold used to 
detect duplications. The proposed patches to address 100\%-similarity duplications 
summed up to +46/-90 (average of +9.2/-18). For a 90\% value, the result was a total 
of +333/-515 with an average of +55.5/-85.83. The 90\% value enabled ArKanjo to identify 
less strict duplication cases, and, therefore, code segments with a greater extension, 
compared to the 100\% similarity entries, which were shorter segments, causing a 
difference in the average number of modified lines.

Finally, the AMD contribution from Group 11, which involves the complete use of the ArKanjo toolchain, 
got a completely different outcome. With a total of 90 lines inserted and 489 lines 
removed, there are some points to be considered for the success and high modification 
extension of the contribution:

\begin{itemize}
    \item The AMD Display subsystem has some self-reported problems regarding overall technical debt;
    \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities.
\end{itemize}

### 2025-07-07 conteudo/tool/architecture.tex
update with the papers updates and revisions

The Function Parser receives the input codebase and transforms it into the temporary codebase. The Function Parser iterates through every source code file from the programming language it supports and uses a specific programming language function extractor to extract every function in the file. For each function extracted, two new source code files and a metadata file are created in the temporary codebase, represented by the pair \textbf{<file name, function name>}, the source code file of the function, and the proper function. The first new source code file contains the function's body from the function it represents, while the other new source code file contains the function's signature. The metadata file contains additional relevant information about the function, such as the function name, the line where the function signature starts in the source code file, and the line where the function's body ends in the source code file. The programming languages supported at the moment are C and Java. Figure \ref{fig:transform} illustrates an example of a function transformation into the temporary codebase.

>>>>

The Function Parser receives the input codebase and transforms it into the temporary
codebase. It iterates through every source code file written in a supported programming 
language and uses a language-specific extractor to isolate each function. For each extracted 
function, two new source code files and a metadata file are created in the temporary codebase, 
represented by the pair \textbf{$<$file name, function name$>$}, the source code of the function, 
and the function metadata (in practice, this is a duplication of the original codebase, 
structured as a parsed representation). One of the new files contains the function body, and the other 
contains the function signature. The metadata file includes additional relevant information,
such as the function name, the line where the signature starts, and the line where the function 
body ends. Currently, the supported programming languages are C and Java, although Java support is limited.
